---
title: "Assignment 2"
author: "Aniket Shah 18200042"
date: "8 November 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Discription
The data set contains seven variables recorded on 71 lawyers in a northeastern American law firm
******
## Task 1 Analysis
**Reading the csv and finding the details of it.**

```{r}
csvData <- read.csv("Lawyers.csv")
inputData <- as.data.frame(csvData, header=TRUE)
dim(inputData)
head(inputData)
```
******
**1. What proportion of the lawyers practices litigation law?**

```{r}
countLitigation <- length(which(inputData$Practice == 'Litigation'))
total <- length(inputData$Practice)
cat('Total numbers of rows', total)
cat('Total Ligitation lawyers count', countLitigation)
pro <- countLitigation / total * 100
cat('Proportion of lawyers practices litigation is: ', format(round(pro, 2)),'%')
```
******
**2. Is the proportion of lawyers in the Boston office that practice corporate law higher than the proportion of lawyers in the Providence office that practice corporate law?**

```{r}
bostonOffice<-nrow(subset(inputData,inputData$Office=="Boston" & inputData$Practice=="Corporate"))
providenceOffice<-nrow(subset(inputData,inputData$Office=="Providence" & inputData$Practice=="Corporate"))

boston<-((bostonOffice/total)*100)
cat('proportion of lawyers in boston office is : ',format(round(boston,2)),'%')

providence<-((providenceOffice/total)*100)
cat('proportion of lawyers in providence office is : ',format(round(providence,2)),'%')

if (boston > providence){
  print('So we can state proportion of practice corporate law is higher in Boston than Providence')
} else {
  print('So we can state proportion of practice corporate law is higher in Providence than Boston')
}
```
******
**3.Use the aggregate function to compute the average age of lawyers who practice corporate law and of lawyers who practice litigation law, across the different levels of seniority. Label the columns of the resulting data frame appropriately. **

```{r}
aggData <- aggregate(inputData$Age~(inputData$Seniority+inputData$Practice), inputData,mean)
colnames(aggData)[1] <- "Seniority"
colnames(aggData)[2] <- "Practice"
colnames(aggData)[3] <- "Average Age"
print(aggData)
```
******
**4. Which office has the youngest median age?**

```{r}
youngData <- aggregate(inputData$Age~inputData$Office, inputData,median)
colnames(youngData)[1] <- "Office"
colnames(youngData)[2] <- "Median Age"
print(youngData)
paste0("Office with minimum median age is: ", youngData[2,1])
```
******
## Task 2: Writing your own function

**1. Write a function which compute the Rosenbrock banana function using a loop. Test the function on the vectors x = (:2; :5) and x = (:2; :5; :1; :6)**

```{r}
# Function using loop to calculate Rosenbrock Banana Calculation
banana_func <- function(x) {
  # Variable to add all individual sum
  summation <- 0
  #To keep the track of iteration
  count <- length(x)
  #iterator to iterate values in given vector
  i <- 1
  #Executing the given function only if length of vector is greater than 1
  if(count > 1){
    # Running loop for calculating function
    while(i <= count-1) {
      # Equation for Rosenbrock banana function
      add <- 100 * ((x[i+1] - x[i]**2)**2) + ((1-x[i])**2)
      summation <- summation + add
      i <- i + 1
    }
  }
  # Returning the final output
  y <- summation
  return(y)
}

x <- c(0.2,0.5)
banana_func(x)

x<-c(0.2,0.5,0.1,0.6)
banana_func(x)
```
******
**2. Propose an alternative function that does not use any loop. Test the function on the same two vectors.**

```{r}
# Function without using loop to calculate Rosenbrock Banana Calculation
banana_func_noloop <- function(xx){
  # Length of the input vector
  d <- length(xx)
  # Creating vector from 1st to second last value
  xi <- xx[1:(d-1)]
  # Creating vector from 2nd to last value
  xnext <- xx[2:d]
  # Sum function to calculate Rosenbrock 
  sum <- sum(100*(xnext-xi^2)^2 + (xi-1)^2)
  y <- sum
  return(y)
}
x <- c(0.2,0.5)
banana_func_noloop(x)

x<-c(0.2,0.5,0.1,0.6)
banana_func_noloop(x)
```
******
**3. Compare the timings you obtain by repeating the function calls 100 times **
```{r}
x = c(0.2, 0.5, 0.1, 0.6)
start_time <- Sys.time()
for(i in 1:100){
  banana_func(x)
}
end_time <- Sys.time()
time_taken_loop <- end_time - start_time
cat('Time taken by function with loop is: ', time_taken_loop, 'seconds')

start_time <- Sys.time()
for(i in 1:100){
  banana_func_noloop(x)
}
end_time <- Sys.time()
time_taken_noloop <- end_time - start_time
cat('Time taken by function without loop is: ', time_taken_noloop, 'seconds')

if (time_taken_noloop > time_taken_loop){
  print('Time taken by no-loop function is more than loop function.')
}else{
  print('Time taken by loop function is more than no-loop function.')
}

```
******
## Task 3: Writing S3 methods

**1. Load in the data as an object called DublinAirport. Assign to the DublinAiport object the classes WeatherData and data.frame.**

```{r}
DublinAirport <- read.csv('2018_09_Dublin_Airport.csv')
class(DublinAirport) <- 'WeatherData'
class(DublinAirport) <- 'data.frame'
class(DublinAirport)
```
******
**2. Write an S3 summary method for an object of class WeatherData which produces the following statistical summaries for the rain, maxtp, mintp variables: mean, standard deviation, minimum, maximum.**
```{r}
summary.WeatherData <- function(wkr) {
  print("Summary for Rain")
  cat("Mean:",mean(data.matrix(wkr[2])))
  cat("\nStd Dev:",sd(data.matrix(wkr[2])))
  cat("\nMinimum:",min(data.matrix(wkr[2])))
  cat("\nMaximum:",max(data.matrix(wkr[2])))
  cat("\n\nSummary for maxtp")
  cat("\nMean:",mean(data.matrix(wkr[3])))
  cat("\nStd Dev:",sd(data.matrix(wkr[3])))
  cat("\nMinimum:",min(data.matrix(wkr[3])))
  cat("\nMaximum:",max(data.matrix(wkr[3])))
  cat("\n\nSummary for mintp")
  cat("\nMean:",mean(data.matrix(wkr[4])))
  cat("\nStd Dev:",sd(data.matrix(wkr[4])))
  cat("\nMinimum:",min(data.matrix(wkr[4])))
  cat("\nMaximum:",max(data.matrix(wkr[4])))
}
summary.WeatherData(DublinAirport)
```
******
**3. Download the new data set 2018 09 Cork Airport.csv from Blackboard, assign the classes WeatherData and data.frame to the object containing the Cork data, and test your function on it. Interpret your findings for Dublin and Cork Airports**

```{r}
CorkAirport <- read.csv('2018_09_Cork_Airport.csv')
class(CorkAirport) <- 'WeatherData'
class(CorkAirport) <- 'data.frame'
class(CorkAirport)

summary.WeatherData(CorkAirport)
```
**Findings:**
Mean rain in Cork is more than Dublin. Additionally, mean mintp in also more in Cork than Dublin. Whereas, mean maxtp is more in Dublin than Cork. 

******
**4. Create an S3 plot method for the class WeatherData that produces the following plots.**  
 ```{r}
 source('params.R')
 ```


```{r}
drawPlot.WeatherData <- function(DublinAirport,tempTitle, rainTitle, maxTempColor, minTempColor){
  library(ggplot2)

firstPlot <- ggplot(DublinAirport, aes(DublinAirport[,1], group = 1)) +
  geom_point(aes(y = DublinAirport[,3], color = maxTempColor)) +
  geom_line(aes(y = DublinAirport[,3], color = maxTempColor)) +
  geom_point(aes(y = DublinAirport[,4], color = minTempColor)) +
  geom_line(aes(y = DublinAirport[,4], color = minTempColor)) +
  labs(x = "Date", y = "Temperature", title = tempTitle) +
  #scale_fill_identity(name = 'the fill', guide = 'legend',labels = c('m1')) +
  scale_colour_manual(name = 'Temparature Label',values =c('Blue'='Blue','Red'='Red'), labels = c('Minimum','Maximum')) +
  theme_classic() +
  geom_vline(aes(xintercept=c(1:30)), linetype="dashed",colour="gray",size=0.7) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
  annotate("text", x = DublinAirport[which.max(DublinAirport[,3])+1,1], y = max(DublinAirport[,3])+2.5, label = "Max Temp") +
  geom_point(data=DublinAirport, aes(x=DublinAirport[which.max(DublinAirport[,3]),1], y=max(DublinAirport[,3])), colour="red", size=3) +
  annotate("text", x = DublinAirport[which.min(DublinAirport[,4])+1,1], y = min(DublinAirport[,4])+2.5, label = "Min Temp")+
  geom_point(data=DublinAirport, aes(x=DublinAirport[which.min(DublinAirport[,4]),1], y=min(DublinAirport[,4])), colour="blue", size=3)


secondPlot <- ggplot(DublinAirport, aes(x=DublinAirport[,1],y = DublinAirport[,2])) +
            geom_bar(stat = "identity") +
              theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) +
              labs(x = "Date", y = "Precipitation Amount (mm)", title = rainTitle) +
              geom_bar(data=data.frame(DublinAirport[which.max(DublinAirport[,2]),1],DublinAirport[which.max(DublinAirport[,2]),2]),aes(DublinAirport[which.max(DublinAirport[,2]),1],DublinAirport[which.max(DublinAirport[,2]),2]),fill=maxRainColor, stat="identity")

library("gridExtra")
grid.arrange(firstPlot, secondPlot)

}
drawPlot.WeatherData(DublinAirport, "Dublin Min- Max Temparature Chart for Sept 2018", "Dublin Rain Bar Chart for Sept 2018", maxTempColor, minTempColor)
drawPlot.WeatherData(CorkAirport,"Cork Min- Max Temparature Chart for Sept 2018", "Cork Rain Bar Chart for Sept 2018", maxTempColor, minTempColor)

```